---
title: "ESM-244-lab-6-MW"
author: "Matthew Warham"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}


library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(ggrepel)
#install.packages("ggspatial")
library(ggspatial)
library(RColorBrewer)
library(raster)


```


Goals: 

- Read in spatial data
- Simplify Polygons
- Add  and transform projections (st_transform, if you know refernece sytem) (st_crs, if you dont know reference system)
- Create maps
- Join spatial data
- Find intersections of spatial data
- Interactive plots leaflet
- Create sf from lat/lon coordinates

###Example 1: Dams in California

```{r}

ca_eco <- read_sf(".", layer = "ca_eco") %>%  # "." gets file from current working directory
  dplyr::select(US_L3NAME) %>% 
  rename(Region = US_L3NAME) %>% 
  st_simplify(dTolerance = 100) %>% #simplifies, lines within 100m will be simplified to one line
  st_transform(crs = 4326)
  

```


```{r set coordinate system for counties data}

ca_counties <- read_sf(".", layer = "california_county_shape_file") 

st_crs(ca_counties) = 4326 # set coordinate system to 4326 (same as ca_eco)

```

```{r}

ca_dams <- read_sf(".", layer = "California_Jurisdictional_Dams") %>% 
  rename(Condition = Condition_) 

ca_dams$Condition <- fct_relevel(ca_dams$Condition, "Fair", "Satisfactory", "Unsatisfactory", "Poor") #Set as factors

```


#MAPS

```{r}

plot(ca_eco)
plot(ca_counties)
plot(ca_dams)

```


#Make a map with ggplot

```{r}

color_count <- 13

my_colors <- colorRampPalette(brewer.pal(10, "Set2"))(color_count) #Make a palette, using all existing colors and then make more on gradient

ggplot(ca_eco)+
  geom_sf(aes(fill = Region),
          color = "NA",  # gets rid of lines
          show.legend = FALSE)+
  scale_fill_manual(values = my_colors)+ # change color scheme to the one we created earlier
  geom_sf(data = ca_counties,
          fill = "NA",
          color = "black", #color of line
          size = 0.1)+
  geom_point(data = ca_dams, # add dams layer
             aes(x = Longitude, y = Latitude),
             size = 1,
             color = "gray10",
             alpha = 0.5)+
  theme_minimal()+
  coord_sf(datum = NA) #remove coordinate axis

```

## Dams in the sierra nevada eco region

```{r}

sn <- ca_eco %>% 
  filter(Region == "Sierra Nevada") %>% # only interested in sierra nevada
  st_join(ca_dams) #join with dams info

ggplot(sn)+ #add this layer because you are going to come back to it for other layers, no need to add 
  geom_sf(data = ca_counties, fill = "wheat3", color = "NA")+ #add counties layer
  geom_sf(fill = "lemonchiffon4", color = "NA")+ #add Sierra nevada layer
  geom_point(aes(x = Longitude, y = Latitude), #only add dams from that are in the sierra nevada
              size = 0.5,
              color = "red4")
  

```

### Example 3: Eco-regions for SB County


```{r}

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara") #four polygons because of the channel islands

eco_clip <- st_intersection(ca_eco, sb) #clip ca_eco by the bounds of the sb polygon


ggplot(eco_clip)+
  geom_sf(data = ca_counties, 
       fill = "grey90", 
       color = "grey80",
       size = 0.2)+
  geom_sf(aes(fill = Region), color = "NA")+
  scale_fill_manual(values = c("darkolivegreen2", "darkolivegreen", "gold2"))+
  coord_sf(xlim = c(-121, -119), ylim = c(33.5,35.5))+ # zoom into desired coordinates
  geom_point(aes(x = -119.6982, y = 34.4208), size = 2)+ # manually create point to plot
  geom_text(x = -119.6982, y = 34.35, label = "I Live Here")+ # put label
  theme_minimal()+
  theme(legend.position = c(0.5, 0.15)) +#move legend
  labs(x = "Longitude", y = "Latitude")

#ggsave

```


###Example 4: Intro to interactive plots with tmap


```{r}

map_sb_eco <- tm_shape(eco_clip)+
  tm_fill("Region", palette = "RdPu", alpha = 0.5)+
  tm_shape(ca_counties)+
  tm_borders()

tmap_mode("view")
map_sb_eco

#leaflet::providers for basemaps
## Example with different basemap
tm_basemap("CartoDB.DarkMatter")+
  tm_shape(eco_clip)+
  tm_borders(col = "white")


```


###Example 5: Fault Lines

```{r}

fault_lines <- read_sf(".", layer = "GMC_str_arc") %>% 
  st_transform(crs = 4326) %>% 
  separate(LTYPE, into =c("syn_ant", "certainty", "direction"), sep = ",") #break information into three categories

plot(fault_lines)

ggplot()+
  geom_sf(data = ca_counties, fill = "black", color = "NA")+
  geom_sf(data = fault_lines, aes(color = syn_ant))+
  theme_void()

# Foult Lines in SB county

sb_faults <- fault_lines %>% 
  st_intersection(sb)

ggplot()+
  geom_sf(data = sb)+
  geom_sf(data = sb_faults, aes(color = syn_ant))

tm_basemap("CartoDB.DarkMatter")+
  tm_shape(sb)+
  tm_borders(col = "grey50", lwd =2)+
  tm_shape(sb_faults)+
  tm_lines(col = "syn_ant", palette = c("orange", "purple"), lwd =2)

```


### Example 6: Faceted Maps

```{r}

ggplot()+
  geom_sf(data = ca_counties, fill = "black", color = "NA")+
  geom_sf(data = fault_lines, aes(color = syn_ant))+
  facet_wrap(~syn_ant)

```


### Example 7: Making spatial points

```{r}

ca_sites <- read_csv("cadfw_sensitive_sites.csv")

View(ca_sites)

sites_sf <- st_as_sf(ca_sites, coords = c("LONDD", "LATDD"), crs = 4326) #converts lat and long columsn into stick geomoetry

View(sites_sf)

ggplot()+
  geom_sf(data = ca_counties, fill = "grey40")+
  geom_sf(data = sites_sf, aes(color = PRI_CODE), size = 0.3)

```


###Example 8: Make Chloropleth of CA counties by number of dams in each county

```{r}

intersection <- st_intersection(ca_dams, ca_counties)

dams_per_county <- intersection %>% 
  group_by(NAME) %>% 
  tally()

dams_per_county


ca_tot <- ca_counties %>% 
  st_join(dams_per_county) %>% 
  dplyr::select(NAME.x,n) %>% 
  rename(name = NAME.x)

#Reassign NA values to zero:
ca_tot$n[is.na(ca_tot$n)] <- 0

ggplot() +
  geom_sf(data =ca_tot, aes(fill = n), size = 0.2) +
  theme_minimal() +
  scale_fill_continuous(low = "yellow", high = "red")


```

